---
title: "HW 3 Task 2 - Cetaceans Raster"
author: "Evan T. Heberlein"
date: "2/22/2021"
output:   
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(raster)
library(here)
library(sf)
library(rnaturalearth)
```

```{r stack rasters, message=FALSE, warning=FALSE}
# Create list of all cetacean rasters
cetacean_rasters <- list.files(
  here("data/ca_cetaceans"),
  full.names = TRUE
)

# Stack all rasters
cetacean_stack <- stack(cetacean_rasters)

# Write function to filter for presence threshold of 0.6
cetacean_filter <- function(x, thresh = .6){
  y <- ifelse(x >= thresh, 1, NA)
  return(y)
}

# Apply filter function to raster stack
cetacean_filtered <- calc(cetacean_stack, fun = cetacean_filter)

# Calculate total species numbers per raster cell
cetacean_sum <- calc(cetacean_filtered, fun = sum, na.rm = TRUE)
```

```{r natural earth, message=FALSE, warning=FALSE, include=FALSE}
# Read in coastal outline from natural earth
pacific_coast <- ne_download(scale = 110, type = 'countries', category = 'cultural', returnclass = 'sf')
```

```{r figure, message=FALSE, warning=FALSE}

# Convert raster to df
cetacean_df <- raster::rasterToPoints(cetacean_sum) %>%
  as.data.frame()

# Create final figure
cetacean_fig <- ggplot() +
  geom_raster(data = cetacean_df, 
                       aes(x = x, y = y, fill = layer)) +
  geom_sf(data = pacific_coast, fill = "black", color = "gray") +
  coord_sf(expand= 0, xlim = c(-125, -115), ylim = c(32, 38)) +
  labs(x = "Longitude", y = "Latitude", fill = "Species richness (n)") +
  theme(legend.position = "bottom") +
  scale_fill_gradient(low = "red", high = "blue")

cetacean_fig
```

***Figure 1** - Cetacean species richness for the Southern California Bight and surrounding areas of the Eastern Pacific Ocean. Species abundance was calculated based on a presence threshold probability of **0.6**, and rasters for all 35 species were aggregated after values over 0.6 were coerced to 1 (and values under 0.6 to 0). Highest species abundance value shown is 28 species (dark blue)*

*Data citation: Kaschner, K., Rius-Barile, J., Kesner-Reyes, K., Garilao, C., Kullander, S., Rees, T., & Froese, R. (2016). AquaMaps: Predicted range maps for aquatic species. www.aquamaps.org*