---
title: "HW 3 Task 3 - War and Peace Text Wrangling"
author: "Evan T. Heberlein"
date: "2/22/2021"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = FALSE)
# Setup - attaching packages
library(tidyverse)
library(tidytext)
library(textdata)
library(pdftools)
library(ggwordcloud)
library(here)
library(patchwork)
```

```{r import, warning=FALSE}

# Read in & tidy text
tolstoy_text <- pdf_text(here("data/war-and-peace.pdf"))

tolstoy_tidy <- data.frame(tolstoy_text) %>% 
  mutate(text_full = str_split(tolstoy_text, pattern = '\\r\n')) %>% 
  unnest(text_full) %>% 
  mutate(text_full = str_trim(text_full))

# Sort data frame by chapter
tolstoy_df <- tolstoy_tidy %>% 
  slice(-(1:12)) %>% 
  mutate(chapter = case_when(
    str_detect(text_full, "Chapter") ~ text_full,
    TRUE ~ NA_character_
  )) %>% 
  fill(chapter) %>% 
  separate(col = chapter, into = c("ch", "no"), sep = " ") %>% 
  mutate(chapter = as.numeric(as.roman(no)))
```

```{r tokens & stop words, message=FALSE, warning=FALSE}

# Isolate words using tokens
tolstoy_tokens <- tolstoy_df %>% 
  unnest_tokens(word, text_full) %>% 
  select(-tolstoy_text) %>% 
  filter(is.na(as.numeric(word)))

# Remove stop words
tolstoy_nonstop_words <- tolstoy_tokens %>% 
  anti_join(stop_words)

# Word count by chapter
nonstop_counts <- tolstoy_nonstop_words %>% 
  count(chapter, word)

# Wordcount for entire book + remove stop words
tolstoy_wordcount_total <- tolstoy_tokens %>% 
  count(word) %>% 
  anti_join(stop_words)

# Top 5 words by chapter - exploration
top_5_words <- nonstop_counts %>% 
  group_by(chapter) %>% 
  arrange(-n) %>% 
  slice(1:5)
```

```{r include=FALSE}
# Download sentiments for `afinn` package
get_sentiments(lexicon = "afinn")
```

```{r sentiment analysis & figures, message=FALSE, warning=FALSE}

# Get top 100 words w/ `afinn` sentiments from whole text
tolstoy_afinn_top100 <- tolstoy_wordcount_total %>% 
  arrange(-n) %>% 
  right_join(get_sentiments("afinn")) %>% 
  slice(1:100)

# Create wordcloud - pt. 1
tolstoy_wordcloud <- ggplot(data = tolstoy_afinn_top100, 
                            aes(label = word)) +
  geom_text_wordcloud(aes(color = value, size = n), 
                      shape = "circle", show.legend = TRUE) +
  scale_color_gradientn(colors = c("red", "blue")) +
  scale_size_area(max_size = 12) +
  labs(color = "Sentiment", size = "Word count") +
  theme_minimal()
  
# Calculate mean sentiment by chapter - unweighted
tolstoy_afinn_ch_avg <- tolstoy_nonstop_words %>% 
  count(chapter, word) %>% 
  right_join(get_sentiments("afinn")) %>% 
  group_by(chapter) %>% 
  summarize(sentiment = mean(value))

# Mean sentiment by chapter weighted by word count
tolstoy_afinn_ch_avg_w <- tolstoy_nonstop_words %>% 
  count(chapter, word) %>% 
  right_join(get_sentiments("afinn")) %>% 
  group_by(chapter) %>% 
  summarize(sentiment = mean(value*n))

# Create weighted & unweighted sentiment graphs by chapter
tolstoy_afinn_bar_weighted <- ggplot() +
  geom_col(data = tolstoy_afinn_ch_avg_w, 
           aes(x = chapter, y = sentiment), fill = "green1") +
  scale_x_continuous(expand = c(0,0), limits = c(0.5, 39), breaks = seq(from = 0,
                                                                      to = 39,
                                                                      by = 3)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.75, 0.75), 
                     breaks = seq(from = -0.75,
                                  to = 0.75,
                                  by= 0.25)) +
  theme_minimal() +
  labs(x = "Chapter", y = "Sentiment (weighted by word count)")

tolstoy_afinn_bar_unweighted <- ggplot() +
  geom_col(data = tolstoy_afinn_ch_avg, 
           aes(x = chapter, y = sentiment), fill = "light blue") +
  scale_x_continuous(expand = c(0,0), limits = c(0.5, 39), breaks = seq(from = 0,
                                                                      to = 39,
                                                                      by = 3)) +
  scale_y_continuous(expand = c(0,0), limits = c(-0.75, 0.75), 
                     breaks = seq(from = -0.75,
                                  to = 0.75,
                                  by= 0.25)) +
  theme_minimal() +
  labs(x = "Chapter", y = "Sentiment (unweighted)")

# Finding 
nonstop_counts_wp <- nonstop_counts %>% 
  filter(word == "war" | word == "peace") %>% 
  mutate(n = as.numeric(n)) %>% 
  mutate(word = as.factor(word))

war_or_peace <- nonstop_counts_wp %>% 
  group_by(chapter) %>% 
  pivot_wider(names_from = word,
               values_from = n) %>% 
  mutate(chapter = as.numeric(chapter))

w_p_plot <- ggplot() +
  geom_line(data = war_or_peace, size = 2,
            aes(x = chapter, y = war, color = "War")) + 
  geom_line(data = war_or_peace, size = 2,
            aes(x = chapter, y = peace, color = "Peace")) +
  theme_minimal() +
  ggtitle("War and Peace in War and Peace") +
  labs(x = "Chapter", y = "Word count", color = "Word") +
  scale_x_continuous(expand = c(0,0), limits = c(1, 39), breaks = seq(from = 0,
                                                                      to = 39,
                                                                      by = 3)) +
  scale_y_continuous(expand = c(0,0)) +
  scale_color_manual(values=c("light blue", "red"))
```

### Pt. 1 - Word counts for entire text of *War and Peace* by Leo Tolstoy

```{r fig 1, message=FALSE, warning=FALSE}
tolstoy_wordcloud
```

***Figure 1** - word cloud showing top 100 words in *War and Peace *displayed by word count and `afinn` sentiment*

### Pt. 2 - Sentiment analysis using `afinn` lexicon

```{r fig 2, message=FALSE, warning=FALSE}
(tolstoy_afinn_bar_weighted | tolstoy_afinn_bar_unweighted)
```

***Figure 2** - mean `afinn` sentiment by chapter of *War and Peace *faceted with `patchwork`. Left-side figure shows sentiment weighted by word count per chapter, and right side shows unweighted sentiment (each `afinn` word that appears in a chapter is weighted equally, regardless of how many times it appears)*

```{r fig 3, message=FALSE, warning=FALSE}
w_p_plot
```

***Figure 3** - bonus figure showing prevalence of words "war" (`afinn` sentiment -2) and "peace" (`afinn` sentiment 2) by chapter in *War and Peace 

Analyzed text citation: *Tolstoy, Leo. War and Peace. Walter Scott Pub. Co., 1920. https://www.planetebook.com/free-ebooks/war-and-peace.pdf*